---
title: "Project_1"
author: "Joseph Elue"
date: "2023-01-16"
output: html_document
---

Section 1 "IO"

Question 1
```{r}
# Suppose we have:
df <- data.frame("y" = rnorm(10, 0, 1), "x" = runif(10, 1, 4))
df

# Save files using write.csv() and write.table()
write.csv(df, file = "df.csv")
write.table(df, file = "df2.csv") 

# Read Files using read.csv() and read.table()
read.csv("df.csv", row.names = 1)
read.table("df2.csv",row.names = 1)
```
Section 2 "VECTORS"

Question 1 

Create each of the following vectors, assign them a “container” and print:
[1] 20 19 18 17 16 15 14 13 12 11 10 9 8 7 6 5 4 3 2 1
[1] 1 2 3 4 5 6 7 8 9 10 11 12 13 14 15 16 17 18 19 20 19 18 17 16 15
[26] 14 13 12 11 10 9 8 7 6 5 4 3 2 1
```{r}
a <- c(20, 19, 18, 17, 16, 15, 14, 13, 12, 11, 10, 9, 8, 7, 6, 5, 4, 3, 2, 1)
b <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 19, 18, 17, 16, 15)
d <- c(14, 13, 12, 11, 10, 9, 8, 7, 6, 5, 4, 3, 2, 1)

print(a)
print(b)
print(d)



```
Question 2

Create the following vector where there are 11 occurrences of 4, 10 occurrences of 6, and 10 occurrences
of 3:
[1] 4 6 3 4 6 3 4 6 3 4 6 3 4 6 3 4 6 3 4 6 3 4 6 3 4 6 3 4 6 3 4
```{r}
a <- c(rep(c(4,6,3),10),4) 
a
```
Question 3

Create a vector of the values of.......
```{r}
a <- c(2,((2^2)/2),((2^3)/3),((2^25)/25))
a
b <- round(a)
b
```
Question 4 

Use paste() to create the following character vectors of length 30: (“label 1”, “label 2”, . . . .., “label
30”). Note that there is a single space between label and the number following.
```{r}
values <- c("label 1", "label 2", "label 30")
second_values <- rep(values, 30)
third_values <- paste(second_values, sep= " ")
third_values
```
Question 5

Execute the following lines which create two vectors of random integers which are chosen with replacement
from the integers 0, 1,. . . , 999. Both vectors have length 250.
set.seed(50)
xVec <- sample(0:999, 250, replace = TRUE)
yVec <- sample(0:999, 250, replace = TRUE)

Code for 5a

```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace = TRUE)
yVec <- sample(0:999, 250, replace = TRUE)

# The index [-1] removes the first y1 in the sequence; thus we can have y2 which would be subtracted from x1
# This goes all the way to yn whihc is the last number 250, while [-length(xVec)] is used because the equation is xn-1
# Which simply means that x(total number of obersations - the last observation) this helps the figures to be balaced considering we already removed # the first value of y

yVec[-1] - xVec[-length(xVec)]

```

Code for 5b 

```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace = TRUE)
yVec <- sample(0:999, 250, replace = TRUE)

# Here indexes were used to pick out the specific points that follow the sin and cosine functions

b <- c(sin(yVec[-length(yVec)]) / cos(xVec[-1]))
b
```
Code for 5c

```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace = TRUE)
yVec <- sample(0:999, 250, replace = TRUE)

d <- c(xVec[1]+2*xVec[2]-xVec[3], 2*xVec[3]-xVec[4],xVec[2-250]+2*xVec[-250]-xVec[250])
d

```
Question 6 

This question uses the vectors xVec and yVec created in the previous question and the functions sort,
order, mean, sqrt, sum and abs.

Code for 6a 

```{r}
set.seed(50)
xVec <- sample(0:999, 250, replace = TRUE)
yVec <- sample(0:999, 250, replace = TRUE)

a <- sqrt(abs(xVec-mean(xVec)))
a
```

Code for 6b 

```{r}
newvector <- yVec[which(yVec > 200)]
a <- length(newvector)
a

```
Code for 6c

```{r}
length(xVec[xVec %% 2 == 0])

```

Code for 6d

```{r}
xVec[order(yVec, decreasing = FALSE)]
```

Code for 6e 

```{r}
newvector <- c(xVec[which(xVec < 300)],xVec[which(xVec > 600)])
newvector
```
 
Section 3 "MATRIXES"

Question 1
Create a 6 × 10 matrix of random integers chosen from 1, 2,. . . , 10 by executing the following two lines
of code:
set.seed(75)
A <- matrix(sample(10, size = 60, replace = TRUE), nrow = 6)

Code for 1a

```{r}
set.seed(75)
A <- matrix(sample(10, size = 60, replace = TRUE), nrow = 6)
A

# Using the colSums function
colSums(A)
rowSums(A)

# Alternative method using the apply function
col.sums <- apply(A, 2, sum)
row.sums <- apply(A, 1, sum)

col.sums
row.sums
```

Code for 1b- f

```{r}
# Find the index of rows whose totals are greater than 50.
set.seed(75)
A <- matrix(sample(10, size = 60, replace = TRUE), nrow = 6)
A
b <- paste(which(rowSums(A) > 50))
b


# Bind those rows in a new matrix (use rbind()).
C <- rbind( A[1,], A[2, ], A[3, ], A[4, ])
C

# Which rows contain exactly two occurrences of the number seven?
d <- which(rowSums(C == 7) == 2)
d

# Find the number of entries in each row which are greater than 4. You can use apply() function.
number_of_rows <- apply(C > 4, 1,  sum)
number_of_rows


# Which rows contain exactly two occurrences of the number seven?
d <- which(rowSums(C == 7) == 2)
d

```
Question 2

```{r}
a <- 1:20
j <- 1:5
solution <- sum(a^4)/(3+j)
sum(solution)

```


Question 3

Russian Roulette is a game played with a revolver and a single bullet. Write an R script which simulates
the game of Russian Roulette, returning “CLICK” 5 out of 6 times, and “BANG” 1 out of 6 times

```{r}
sample (c(" BANG ", " click ") , size =1 , prob =c(1/6 , 5/6) )
```
Question 4 

```{r}
# Question 4a
M <- matrix(sample(c(1:10), 16, replace = TRUE), nrow = 4, byrow = TRUE)
M

max_value <- which(M == max(M), arr.ind=TRUE)
max_value

min_value <- which(M == min(M), arr.ind=TRUE)
min_value

# Question 4b

#max_value2 <- apply(which(M[] == max(M), arr.ind=TRUE)) 
#max_value2


# Convert M to a list of column-vectors. Look at split()
a <- split(M, rep(1:ncol(M), each = nrow(M)))

print(a)

```

Section 4 "FUNCTIONS"

Question 1 and 2 

```{r}
# Code 
x <- c(1:10)

fun1 <- function(fun1){
  n <- length(x)
  return(fun1^(1:n))
}

y <- fun1(x)
y

# Code 
fun2 <- function(fun2){
  n <- length(x)
  return(fun2^(1:n)/(1:n))
}
z <- fun2(x)
z

# Try out 
fun3 <- c(1:5,6:1)
fun1 <- function(fun1){
  n <- length(fun3)
  return(fun3^(1:n))
}

b <- fun1(fun3)
b


```
Question 3

```{r}
x <- c(1:10)

fun5<- function(x)
{
ifelse(x < 0, x^2 + 2*x + 3, ifelse(x < 2, x+3, x^2 + 4*x - 7))
}

graph <- seq(-3:3)
plot(graph, fun5(graph), type="l")
```

Question 4 

```{r}
X <-  matrix(c(1,1,3,5,2,6,-2,-1,-3), nrow = 3, byrow = TRUE)
X

fun6 <- function(X) {
  ifelse(X %% 2 != 0, 2*X, X)
}
w <- fun6(X)
w

```

Question 5

```{r}
set.seed(3)
M <- matrix(sample(c(1:8), 16, replace = TRUE), nrow = 4, byrow = TRUE)
M

getmode <- function(v) {
a <- table(v, v)
b <- diag(a)
return(as.numeric(names(which.max(b))))
}
getmode(M)


# apply(getmode(M), 2,  sum)

```


Section 5 "Conditions"

Question 1

Take the sqrt() of the following vector by skipping (-) numbers. Use ifelse() first. It will give your
some NAs. Now use if/else conditions openly to avoid these NAs

```{r}
set.seed(2)
d <- sample(-10:10, 10, replace = TRUE)
d

# Take the sqrt() of the following vector by skipping (-) numbers. Use ifelse() first
b <- ifelse(d > 0, sqrt(d), NA)
b

# Now use if/else conditions openly to avoid these NAs
if(anyNA(b)){
print(b[!is.na(b)])
} else {
print("No NA")
}

```
Question 2 

Now use the same vector, d, and apply sqrt() if the number is larger than or equal to 0, if not take
the square of the number

```{r}
set.seed(2)
d <- sample(-10:10, 10, replace = TRUE)
d

b <- ifelse(d >= 0, sqrt(d), d^2)
b
```
Question 3.

You have the following matrix. Create a code that prints how many numbers there are less than 6 in
A[1,]. If there is no number less than 6 print “No number less than 6 in Row 1”. Make sure that you
print only once! Hint use any() in the condition. Use if and else separately instead of ifelse().

```{r}
set.seed(1)
A <- matrix(sample(1:10, 25, replace = TRUE), 5, 5)
A

# Create a code that prints how many numbers there are less than 6 in A[1,]
a <- c((A[1, ]< 6))
sum(a)

ifelse(sum(c((A[1, ]< 6))), "Numbers less than 6 in Row 1", "No number less than 6 in Row 1" )

```
Section 6 "LOOPS"

Question 1

Create a loop that runs 100 times. In each run, it draws 1000 numbers from a standard normal
distribution. It calculates the mean and the standard deviation of these 1000 numbers. Finally, it saves
them to separate containers, one for means and one for standard deviations.

```{r}
x_mean <- c()
x_sd <- c()

for (i in 1:100) {
  set.seed(i)
  x <- rnorm(1000)
  x_mean[i] <- mean(x)
  x_sd[i] <- sd(x)
}
```

Question 2

The numerical or character variable could be an indicator variable. For example, if a numerical variable
has less than a certain threshold (like 10 unique numbers), it might be considered an indicator variable.
Write a code that looks each variable in mtcar (see datasets package) and identifies each indicator
variable and convert it to a factor variable in a loop.


```{r}
str(mtcars)

x_mpg <- c()
x_cyl <- c()
x_disp <- c()
x_hp <- c()
x_drat <- c()
x_wt <- c()
x_qsec <- c()
x_vs <- c()
x_am <- c()
x_gear <- c()
x_carb <- c()

for (i in mtcars) {
  set.seed(i)
  mpg <- c(mtcars$mpg)
  x_mpg <- as.factor(mpg)
  
}
for (i in mtcars) {
  set.seed(i)
  cyl <- c(mtcars$cyl)
  x_cyl <- as.factor(cyl)
}
for (i in mtcars) {
  set.seed(i)
  disp <- c(mtcars$disp)
  x_disp <- factor(disp)
}
for (i in mtcars) {
  set.seed(i)
  hp <- c(mtcars$hp)
  x_hp <- factor(hp)
}
for (i in mtcars) {
  set.seed(i)
  drat <- c(mtcars$drat)
  x_drat <- factor(drat)
}
for (i in mtcars) {
  set.seed(i)
  wt <- c(mtcars$wt)
  x_wt <- factor(wt)
}
for (i in mtcars) {
  set.seed(i)
  qsec <- c(mtcars$qsec)
  x_qsec <- factor(qsec)
}
for (i in mtcars) {
  set.seed(i)
  vs <- c(mtcars$vs)
  x_vs <- factor(vs)
}
for (i in mtcars) {
  set.seed(i)
  am <- c(mtcars$am)
  x_am <- factor(am)
}
for (i in mtcars) {
  set.seed(i)
  gear <- c(mtcars$gear)
  x_gear <- factor(gear)
}
for (i in mtcars) {
  set.seed(i)
  carb <- c(mtcars$carb)
  x_carb <- factor(carb)
}

class(x_mpg)
class(x_cyl)
class(x_disp)
class(x_hp)
class(x_drat)
class(x_wt)
class(x_qsec)
class(x_vs)
class(x_am)
class(x_gear)
class(x_carb)

```







